---
title: "Homework 10"
author: "Key"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!    

## Background
In the `data` folder, you will find data about shark incidents in California between 1950-2022. The [data](https://catalog.data.gov/dataset/shark-incident-database-california-56167) are from: State of California- Shark Incident Database.   

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the data
Run the following code chunk to import the data. In this chunk, I am removing NA's.
```{r message=FALSE, warning=FALSE}
sharks <- 
  read_csv("data/SharkIncidents_1950_2022_220302.csv", 
           na=c("NA", "N/A", "Unknown", "NOT COUNTED", "", " ")) %>% 
  clean_names()
```

## Questions
1. Are there any "hotspots" for shark incidents in California? Show the top five counties with the most incidents.

```{r}
sharks %>% 
  count(county) %>% 
  slice_max(n, n=5) 
#San Diego and Humboldt counties seem to be hotspots as they are the only 2 places that break the 20 incident count
```


2.  Make a plot that shows the total number of incidents per county.
```{r}
sharks %>% 
  count(county) %>% 
  slice_max(n, n=5) %>% #assuming graph is wanted for only top 5 counties, if not then I wouldve done same but without slice_max
  ggplot(mapping=aes(y=n, x=county, fill=county))+
  geom_col()+
  labs(title = "Total Number of incidents per County",
       y="Number of Incidents",
       x="County")+
  coord_flip()
```

3. Are there months of the year when incidents are more likely to occur? Make a plot that shows the total number of incidents by month. Which month has the highest number of incidents?  
```{r}
sharks %>% 
  count(month)
```

```{r}
sharks %>% 
  mutate(month=as.factor(month)) %>% 
  count(month) %>%
  ggplot(aes(y=n, x=month, fill=month))+
  geom_col()+
  labs(title="Number of Incidents Every Month",
       x="Month by its Date Number",
       y="Number of Incidents")
```

4. Is there a time of day when most attacks occur? Make a plot that shows the total number of incidents by time of day. Which time of day has the highest number of incidents?  
```{r}
sharks %>% 
  count(time) %>% 
  slice_max(n, n=10)
```


```{r}
sharks %>% 
  count(time) %>%
   slice_max(n, n=10) %>% #Too many points, so reduced to most relevant
  ggplot(aes(x=time, y=n, fill = time))+
  geom_col()
#11:00 am stands out as the time period with a noticeable spike in Shark attacks 
```


5. How do the number and types of injuries compare by county? Make a table that shows the number of injury types by county. Which county has the highest number incidents? (hint: your table should be in wide format)

```{r}
sharks %>% 
  count(county, injury) %>% 
  pivot_wider(names_from = "injury",
              values_from = "n")
#Humboldt County has the highest number of incidents with 20 recorded incidents.
```


6. Use the table from #5 to make a plot that shows the total number of incidents by county.
```{r}
sharks %>% 
  ggplot(aes(x=county, y=injury, fill = injury))+
  geom_col(position = "dodge")+
  coord_flip()+
  labs(title = "Total Number of Incidents by County")
```

7. In the data, `mode` refers to a type of activity. Which activity is associated with the highest number of incidents?
```{r}
sharks %>% 
  count(mode)
#Surfing/Boarding is associated with the highest amount of incidents
```

8. Make a plot that compares the number of incidents by activity.
```{r}
sharks %>%
  ggplot(aes(x=mode, fill = mode))+
  geom_bar(position = "dodge")+
  coord_flip()+
  labs(title = "Number of Incidents by Activity",
       x="Activity")
```

9. Which shark species is involved in the highest number of incidents?  
```{r}
sharks %>% 
  count(species)
#Great White Sharks make up the overwhelming majority of incidents
```

10. Are all incidents involving Great White's fatal? Make a plot that shows the number and types of incidents for Great White's only.  
```{r}
sharks %>% 
  filter(species=="White") %>% 
  ggplot(aes(x=species, fill = injury))+
  geom_bar(position = "dodge")+
  labs(title = "Number of Each Type of Incident for Great White Sharks",
       x="Injury Type")
```

## Knit and Upload
Please knit your work as an .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! 